---
title: "Analytics design-HW3"
output: pdf_document
date: "2/9/2020"
author: MSBA Team11
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load library and datasets

```{r}
rm(list = ls())
library(cluster)
library(fpc)
require("cluster")
require("fpc")
require("factoextra")
require("gridExtra")
library(fpc)
library(factoextra)
library(gridExtra)
library('reshape')
load(file='GBA424 - Toy Horse Case Data.Rdata')
```

########## Part A ##############
A. Use Regression to estimate the conjoint model at the individual level
• Produce part-utilities to pass to part B
• Produce predictions for missing profiles (4 profiles) to pass to part D

Each parent rated a total of 12 profiles. 
total:16 profiles
```{r}
# divide the conjontData into two parts: one for part-utilities, and another for prediction
predictData  = conjointData[is.na(conjointData$ratings),]
regData = conjointData[!is.na(conjointData$ratings),]

## part-utilities & predictions
df = c()
for (i in unique(regData$ID)){
  data = regData[regData$ID == i,]
  pre.Data = predictData[predictData$ID == i,]
  model = lm(ratings ~ price + size + motion + style,data)
  df = c(df,model[['coefficients']])
  profile.pre = predict(model,pre.Data)
  for (m in 1:length(profile.pre)){
      predictData[predictData$ID==i,][m,]$ratings=profile.pre[[m]]
  }
}
dim(df) = c(5,200)
df = t(df)
df = data.frame(df)
colnames(df) = c('Intercept','price-119.99','size-26inches','motion-Rocking','style-Glamour')

# part-utilities
part.utilities = df
# after adding the predictions
conjont.Data = rbind(regData,predictData)
conjont.Data = conjont.Data[order(conjont.Data$ID,conjont.Data$profile),]

```

########## Part B ##############
B. Conduct Benefit Segmentation via Cluster Analysis of Conjoint Part-Utilities.
• Use cluster analysis on the part-utilities (including the constant)
• Test at least two cluster analysis schemes (i.e., number of clusters) and select the
best one in your view. Justify this decision.
• Interpret the segments in the chosen scheme and identify the ideal product for
each segment

```{r}
########Evaluate number of clusters to use on data with visualizations
clustTest = function(toClust,print=TRUE,scale=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100){
  if(scale){ toClust = scale(toClust);}
  set.seed(seed);   # set random number seed before doing cluster analysis
  wss <- (nrow(toClust)-1)*sum(apply(toClust,2,var))
  for (i in 2:maxClusts) wss[i] <- sum(kmeans(toClust,centers=i,nstart=nstart,iter.max=iter.max)$withinss)
  ##gpw essentially does the following plot using wss above. 
  #plot(1:maxClusts, wss, type="b", xlab="Number of Clusters",ylab="Within groups sum of squares")
  gpw = fviz_nbclust(toClust,kmeans,method="wss",iter.max=iter.max,nstart=nstart,k.max=maxClusts) #alternative way to get wss elbow chart.
  pm1 = pamk(toClust,scaling=TRUE)
  ## pm1$nc indicates the optimal number of clusters based on 
  ## lowest average silhoutte score (a measure of quality of clustering)
  #alternative way that presents it visually as well.
  gps = fviz_nbclust(toClust,kmeans,method="silhouette",iter.max=iter.max,nstart=nstart,k.max=maxClusts) 
  if(print){
    grid.arrange(gpw,gps, nrow = 1)
  }
  list(wss=wss,pm1=pm1$nc,gpw=gpw,gps=gps)
}

clustTest(part.utilities) # cluster = 3
```
```{r}
########Runs a set of clusters as kmeans
runClusts = function(toClust,nClusts,print=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100){
  if(length(nClusts)>4){
    warning("Using only first 4 elements of nClusts.")
  }
  kms=list(); ps=list();
  for(i in 1:length(nClusts)){
    kms[[i]] = kmeans(toClust,nClusts[i],iter.max = iter.max, nstart=nstart)
    ps[[i]] = fviz_cluster(kms[[i]], geom = "point", data = toClust) + ggtitle(paste("k =",nClusts[i]))
   
  }
  library(gridExtra)
  if(print){
    tmp = marrangeGrob(ps, nrow = 2,ncol=2)
    print(tmp)
  }
  list(kms=kms,ps=ps)
}
cluster.selection=runClusts(part.utilities,c(3,4,5)) # according to the graphs, cluster =3 is the best option
cluster_3 = cluster.selection$kms[[1]]
cluster_3 #result
```

```{r}
########Plots a kmeans cluster as three plot report
plotClust = function(km,toClust,discPlot=FALSE){
  nc = length(km$size)
  if(discPlot){par(mfrow=c(2,2))}
  else {par(mfrow=c(3,1))}
  percsize = paste(1:nc," = ",format(km$size/sum(km$size)*100,digits=2),"%",sep="")
  pie(km$size,labels=percsize,col=1:nc)
  
  clusplot(toClust, km$cluster, color=TRUE, shade=TRUE,
           labels=2, lines=0,col.clus=1:nc); #plot clusters against principal components
  
  if(discPlot){
    plotcluster(toClust, km$cluster,col=km$cluster); #plot against discriminant functions ()
  }
  rng = range(km$centers)
  dist = rng[2]-rng[1]
  locs = km$centers+.05*dist*ifelse(km$centers>0,1,-1)
  bm = barplot(km$centers,beside=TRUE,col=1:nc,main="Cluster Means",ylim=rng+dist*c(-.1,.1))
  text(bm,locs,formatC(km$centers,format="f",digits=1))
}
plotClust(cluster_3,part.utilities)
```

########## Part C ##############
C. Conduct a priori segmentation.
• Conduct a priori segmentation using the variables gender and age in order to
profile the attribute preferences based on these variables (use segment-level
regressions)
• Test whether these a priori segmentation variables affect the part-utilities. What
does this test tell you about these as segmentation schemes?
• If the differences are meaningful, profile the segment-level attribute preferences
and identify the ideal product for the a priori segments.

```{r}
survData = merge(regData,respondentData,by='ID')

reg_gender = lm(ratings ~ (price + size + motion + style):gender,survData)
summary(reg_gender)

reg_age = lm(ratings ~ (price + size + motion + style):age,survData)
summary(reg_age)
##the differences are meaningful
```
```{r}
## a priori segmentation:gender
gender_0M = survData[survData$gender == 0,]
reg_genderM = lm(ratings ~ price + size + motion + style,gender_0M)
# reg_genderM$coefficients
summary(reg_genderM)

gender_1F = survData[survData$gender == 1,]
reg_genderF = lm(ratings ~ price + size + motion + style,gender_1F)
# reg_genderF$coefficients
summary(reg_genderF)
```


```{r}
## a priori segmentation:age
age_0.2 = survData[survData$age == 0,]
reg_age_2 = lm(ratings ~ price + size + motion + style,age_0.2)
# reg_age_2$coefficients
summary(reg_age_2)


age_1.34 = survData[survData$age == 1,]
reg_age_34 = lm(ratings ~ price + size + motion + style,age_1.34)
# reg_age_34$coefficients
summary(reg_age_34)
```

########## Part D ##############
D. Simulate market shares for different product-line scenarios.
• Use disaggregate analysis with a first choice rule to forecast market shares. Using
these market shares and the information about costs in the case, calculate
profitability for each product in the product line as well as the overall
profitability for the firm.
• You should present at least 4 scenarios that try to identify the best possible
product line strategy given considerations related to competitive response,
cannibalization, profitability, and long-run performance.
• Be sure to briefly justify why you chose the scenarios you chose!


total annual market size is 4,000 units
variable cost:
18" Rocking $33
26" Rocking $41
18" Bouncing $21
26" Bouncing $29

Costs = 20,000/year * #products + $20,000/3 *#products not in existing set
```{r}
ratings = conjont.Data[,1:3]
ratings = cast(ratings,formula =ID~profile)

simFCShares = function(scen,data,ascend=FALSE){
  decs = data.frame()
  inmkt = data[,scen] #construct the subsetted matrix of options
  if(ascend){ #if ranks 1 is best
    bestOpts = apply(inmkt,1,which.min)  #identify which option is best = min
  } else { #else the best rank is the largest number
    bestOpts = apply(inmkt,1,which.max) #identify which option is best = max
  }
  for (i in 1:nrow(inmkt)){
      for (m in 1:length(inmkt[i,])){
          if (inmkt[i,m] == inmkt[i,bestOpts[i]]){
             decs[i,m] = 1
          }else{
              decs[i,m] = 0
          }
      }
      if (rowSums(decs)[i]>1){
          decs[i,][decs[i,]!=0] = 1/rowSums(decs)[i]
      }
  }

  # decs = as.data.frame(model.matrix(~0+as.factor(bestOpts))) #fill to set of options marked 0 or 1
  shs = colSums(decs)/sum(decs) #assumes that total decisions is market size
  names(shs) = names(inmkt) #attach labels
  shs
}

# colmeans = colMeans(ratings) 
# age = prop.table(table(respondentData$age))
# age
# gender = prop.table(table(respondentData$gender))
# gender

```

```{r}
###Set up scenarios
# 7 is the competitor's product
# 5 and 13 are the current products
# cluster prefer proflie:4 14 16; 3 clusters of sizes 80, 53, 67
# age-2:16, age3-4:12, male:16, female:4

scens = list()
scens[[1]]=c(5,7,13) # current market
scens[[2]]= c(5,7,13,16)       
scens[[3]]= c(5,7,13,4) 
scens[[4]]= c(5,7,12,13)
scens[[5]]= c(5,7,13,14)
scens[[6]]= c(5,7,16)
scens[[7]]= c(5,8,13,16)       
scens[[8]]= c(5,8,13,4) 
scens[[9]]= c(5,8,12,13)
scens[[10]]= c(5,8,13,14)
scens[[11]]= c(5,8,16)
scens[[12]]= c(5,8,4)
scens[[13]]= c(8,14,16)
scens[[14]]= c(8,4,16)
scens[[15]]= c(8,4,14,16)
scens[[16]]= c(8,4,12,16)
scens[[17]]= c(8,4,12)
scens[[18]]= c(8,12,16)
scens[[19]]= c(8,12,14)
scens[[20]]= c(8,4,12,14)

for (i in 1:length(scens)){
    n = simFCShares(scens[[i]],ratings[,2:17])
    print(n)
}

```

